### 2014-04-07 presentation/graphic/time_line.tex
Presentation finished up

\documentclass[10pt, varwidth]{standalone}

%
% compile using: pdflatex -shell-escape reu_paper.tex
%

\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{shapes,arrows}

\usepackage{pgfplots}

\usepackage{caption}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{mdwlist}
\usepackage{mathtools}
\usepackage{url}

\date{}

\begin{document}
\maketitle
\vspace{-3cm}

\begin{tikzpicture} [scale=.5]
		\draw [thick] (1.5,6.5) -- (13.5,6.5) -- (13.5,3.5) -- (1.5,3.5) -- (1.5,6.5);
		\node[align=center] at (4.5,5) {\textbf{Tiles to assign:}};
		\node[align=center, box, draw] at (8.5,5) {13\\9};
		\node[align=center, box, draw] at (10.5,5) {11\\13};
		\node[align=center, box, draw] at (12.5,5) {9\\12};		
		\draw [thick] (-4,2) -- (18,2);

		\node[align=right] at (11,2.5) {\textbf{Tiles}};

		\node[align=center, box, draw] at (5,0) {13\\9};
		\node[align=center] at (6.5,0) {1.5};
		\node[align=center, box, draw, fill = green] at (11,0) {11\\13};
		\node[align=center] at (12.5,0) {0.5};
		\node[align=center, box, draw, fill = orange] at (16.5,0) {9\\12};
		\node[align=center] at (18,0) {2.5};

		\draw (11,-1) -- (9.5,-1.5);
		\draw (11,-1) -- (13,-1.5);

		\node[align=center, box, draw, fill = green] at (9.5,-2.5) {13\\9};
		\node[align=center] at (11,-2.5) {1.0};
		\node[align=center, box, draw, fill = orange] at (12.5,-2.5) {9\\12};
		\node[align=center] at (14,-2.5) {8.0};

		\draw (5,-1) -- (3.5,-1.5);
		\draw (5,-1) -- (6.5,-1.5);

		\node[align=center, box, draw, fill = orange] at (3.5,-2.5) {13\\9};
		\node[align=center] at (5,-2.5) {6.0};
		\node[align=center, box, draw, fill = orange] at (6.5,-2.5) {9\\12};
		\node[align=center] at (8,-2.5) {5.0};

		\draw (9.5,-4) -- (9.5,-3.5);

		\node[align=center, box, draw, fill = green] at (9.5,-5) {9\\12};
		\node[align=center] at (11,-5) {1.6};

		\draw [thick] (2,3) -- (2,-6);

		\node[align=right] at (-1,2.5) {\textbf{Protein Chain}};
		\node[align=right] at (-1,0) {11.5};
		\node[align=right] at (-1,-2.5) {12.5};
		\node[align=right] at (-1,-5) {9.6};

	\end{tikzpicture}
\end{document}