### 2014-04-08 presentation/percent_graph/test.tex

% to run use
% $ pdflatex -shell-escape plot.tex

\documentclass[varwidth]{standalone}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{graphics}
\usepackage{graphicx}

\begin{document}
\begin{figure}
\begin{center}
\begin{tikzpicture}

\begin{axis}[width = \columnwidth, grid=major, xlabel={Number of Residues}, ylabel={Time to Completion (seconds)}]
\addplot[no markers, red] gnuplot [raw gnuplot] { 
fac(x) = (int(x)==0) ? 1.0 : int(x) * fac(int(x)-1.0);
g(x) = (a*x**2+b*x+c)/(d*x+e);
fit g(x) 'percent.dat' using 1:(log10($2/fac($1))) via a,b,c,d,e;
plot 'percent.dat' using 1:(log10($2/fac($1))) title 'Data', g(x);
};

\end{axis}
\end{tikzpicture}
% \caption{Time to completion for analysis of a uniform filesystem with uniform branching factor two plotted with entry counts.}
% \label{fig:radix_2_line}
\end{center}
\end{figure}
\end{document}